From physical principles to classical Hamiltonian mechanics

### 2014-09-13 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
First pass in using set notation for labels

\section{States}
\textbf{State definition}. We assume we have a physical system that can be found in different configurations. We call each unique configuration \emph{configuration state} and assign it a \emph{label}. For our purposes, it does not matter what the label is (a measurement, a way to prepare the system, future behavior, ...), just that it exists.

\begin{defn}\label{statedef}
Let $I$ be a set of labels. Let $\mathbbm{C}_{I}$ be the set of all possible configuration states of our system identified by the labels. $\forall i \in I, \exists \mathbbm{c}_{i} \in \mathbbm{C}_{I}$.
\end{defn}

The following definitions allow us to use multiple labels.

\begin{defn}\label{labelsCombine}
Let $I_1$ and $I_2$ be two sets of labels. We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map $f:I \rightarrowtail I_1, I_2$ exists that maps an element of $I$ to a pair of elements of $I_1$ and $I_2$.
\end{defn}

\begin{defn}\label{labelsOrth}
The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
\end{defn}

The following assumptions and definitions allow us to describe a configuration state by its parts.

@@

\begin{defn}\label{classicalPhaseSpace}
Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
\end{defn}

\begin{cor}\label{classicalDistribution}
Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$. In physics term, the distribution can be visualized as a histogram over the discrete labels.
\end{cor}

Under the classical assumption, we can then limit ourselves to study the labels of states and their properties without loss of generality.

@@

\begin{prop}\label{detrevMap}
Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
\end{prop}

\begin{cor}\label{detrevDist}
The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
\end{cor}

@@

\begin{defn}\label{labelsIndep}
Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$. Let $f:I \leftrightarrow I$ be a bijective map. We call the labels \emph{independent} iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
\begin{enumerate}
\item $I_1' \times I_2' = I$
\item there exists a bijective map $f_1: I_1 \leftrightarrow I_1'$
\item there exists a bijective map $f_2: I_2 \leftrightarrow I_2'$
\item $f(i) = \langle f_1(i_1), f_2(i_2)\rangle$
\end{enumerate}
\end{defn}

>>>>

\section{States, Labels and Maps}

\begin{defn}\label{statedef}
Fix a physical system for study. We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system. Each element $\mathbbm{c}$ we call \emph{configuration state}.
\end{defn}

@@

\begin{defn}\label{classicalPhaseSpace}
Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}$ a \emph{state}.
\end{defn}

\begin{cor}\label{classicalDistribution}
Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$. The distribution can be visualized as a histogram over the states in phase space.
\end{cor}

Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.

\begin{defn}\label{label}
To help identify states, we introduce the following: we call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$. Therefore a state is in one and only one label of a state variable.
\end{defn}

\begin{defn}\label{labelsCombine}
Let $I_1$ and $I_2$ be two sets of labels. We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$. If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{orthogonal}, and the total number of labels of $\langle I_1, I_2 \rangle$ is given by the product of the number of labels in $I_1$ and $I_2$.
\end{defn}

@@

\begin{prop}\label{detrevMap}
Let $S$ be the set of configuration states of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:S \leftrightarrow S$.
\end{prop}

\begin{cor}\label{detrevDist}
The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
\end{cor}

@@

\begin{cor}\label{detrevDist}
Given a label $i$, the image $f(i)$ is also a label containing the same number of states. Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels. Given a state variable $X$, the image $f(X)$ is also a state variable.
\end{cor}

\begin{defn}\label{labelsIndep}
Let $I_1$ and $I_2$ be two orthogonal sets of labels. They are said to be \emph{independent} if and only if the image of the combined set $f(\langle I_1, I_2 \rangle)$ is equal to the combined set of the images $\langle f(I_1), f(I_2) \rangle$. It follows that $f(I_1)$ and $f(I_2)$ are orthogonal.
\end{defn}
